Combining information from photometric and spectroscopic data:\\ Don't waste your time in spectrophotometric calibration!

### 2014-04-26 tex/ms.tex
equations, notes to self

\documentclass[iop,numberedappendix]{emulateapj}
\usepackage{apjfonts}
\bibliographystyle{apj}

\input{vc}

@@

Hello World.

@@

Hello World.

@@

>>>>

% \documentclass[iop,numberedappendix]{emulateapj}
% \usepackage{apjfonts}
\documentclass[12pt, letterpaper, preprint]{aastex}
\bibliographystyle{apj}

\input{vc}
\newcommand{\transpose}[1]{{#1}^{\!\mathsf T}}
\newcommand{\given}{\,|\,}

@@

THIS SECTION NEEDS WORK; THIS IS JUST AN EQUATION DUMP.

The standard chi-squared or log-Gaussian log-likelihood is
\begin{eqnarray}
\ln p(d\given\theta) &=& -\frac{1}{2}\,\sum_{n=1}^N \left[\frac{[d_n - \mu_n(\theta)]^2}{\sigma_n^2} + \ln(2\,\pi\,\sigma_n^2) \right]
\quad ,
\end{eqnarray}
where $\theta$ is a blob of parameters of interest,
$d$ is the full set (in the form of a column vector) of data,
made up of $N$ values $d_n$,
$\mu(\theta)$ is the the column-vector mean prediction (expectation value) of the model,
made up of $N$ values $\mu_n(\theta)$,
the $\sigma_n^2$ are the observational noise variances.
We have included the $\ln(\sigma_n^2)$ term in the likelihood because
in what follows we are going to be modifying the noise model.

The calibration-vector and Gaussian-Process modified log-likelihood is
\begin{eqnarray}
\ln p(d\given\theta,\alpha) &=& -\frac{1}{2}\,\left[\transpose{\Delta}\,C^{-1}\,\Delta + \ln([2\,\pi]^N\,\det{C}) \right]
\\
\Delta &\equiv& d - f(\alpha)\ast\mu(\theta)
\\
C_{nn'} &\equiv& \sigma_n^2\,\delta_{nn'} + K_\alpha(\lambda_n - \lambda_n')
\quad ,
\end{eqnarray}
where $\alpha$ is a blob of nuisance parameters
(which determine the calibration vector and the kernel function),
$\Delta$ is a residual column vector,
$f(\alpha)$ is a calibration column vector,
the $\ast$ operator is the pixel-wise multiplication operator,
$C$ is the Gaussian Process covariance matrix,
made up of $N^2$ values $C_{nn'}$,
$K_\alpha$ is a kernel function,
and each $\lambda_n$ is the wavelength of spectral pixel $n$.
and $a^2$ and $\ell^2$ are parameters of the kernel function (and hence elements of $\alpha$).
The calibration vector $f(\alpha)$ takes care of the large-scale
calibration residual; the Gaussian process is going to pick up
smaller-scale residuals away from \emph{that}.

In detail, we are going to choose a $M$th-order polynomial form for
the calibration vector $f(\alpha)$ and a bell-shape (Gaussian) for the
kernel function $K_\alpha(\cdot)$:
\begin{eqnarray}
f(\alpha) &=& \sum_{m=1}^M c_m\,x^m
\\
x &\equiv& \frac{\lambda}{\lambda_0} - 1
\\
K_\alpha(x) &=& a^2\,\exp(-\frac{x^2}{2\,\ell^2})
\\
\transpose{\alpha} &\equiv& \left[ \{c_m\}_{m=0}^M, a^2, \ell^2 \right]
\quad ,
\end{eqnarray}
where $\lambda$ is a column vector made up of the wavelengths $\lambda_m$
and $x$ is a scaled and shifted version of the same,
scaled by a fiducial wavelength $\lambda_0$.
Nuisance parameters $a^2$ and $\ell^2$ set the variance and
correlation length of the covariant calibration noise, respectively.

@@

BDJ INSERT WORDS HERE.

@@

BDJ INSERT WORDS HERE.

HOGG COMPUTE TOTAL ANNUAL COST OF SPECTROPHOTOMETRY (to astronomical community).
